\MakeUppercase{The Structure of $k$-Planar Graphs} \thanks{This work was partly funded by NSERC and MRI.}

### 2019-07-04 k-planar.tex
Pffftt

For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.  The \emph{$(k,d)$th power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{d,k})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.

Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} ot $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.  For simplicity, we bypass the graph $H'$ and assume, from the outset, that $V(H)\subseteq V(T)$.

For two nodes $vw\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.  Observe that every $(k,d)$ $vw$-shortcut is also a $(k,d)$ $vw$-longcut and that $G$ contains a $(k,d)$ $vw$-shortcut if and only if it contains a $(k,d)$ $vw$-longcut.

For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $P$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $P$.  We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $G$ contains a $vw$-longcut that crosses $x$.

\begin{clm}
  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node in which $x$ participates. 
\end{clm}

@@

  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.  The connectivity of $G[X]$ implies that $H[X_H]$ is connected.  Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.  Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
\end{proof}

For each $x\in V(T)$, we define the \emph{separator}
\[  
     S_x := \{v\in V(G): a(v)= x\} \enspace .
\]
Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.  We let $I:=G^{k,d}/(S_x : x\in V(T))$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$. (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)

@@

  By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.  Therefore, $G$ contains a path $v=v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.  Since $v\not\in V_x$ and $v_r\in V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$. Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some srict $T$-ancestor $a$ of $x$.  But the existence of the path $v_0,\ldots,v_i$ in $G$ implies that $v=v_0\in S_{a'}$ for some ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
\end{proof}

@@

\begin{clm}\clmlabel{general-width}
  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le t\ell(1 + 2(k-1)d^{k-1})$.  
\end{clm}

\begin{proof}
  Consider some vertex $v\in S_x\cap L_i$.  There are two cases to consider:
  \begin{enumerate}
    \item $v\in Y_x$.  By \yref{thickness}, there are at most $t\ell$ such vertices in $S_x\cap L_i$.
    \item $G$ contains a path $v=v_0,\ldots,v_r$ with $r<k$ such that $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$.
    In this case, $v_r\in L_{i+j}\cap Y_x$ for some $j\in\{-k+1,\ldots,k-1\}$.
    For each $w\in L_{i+j}\cap Y_x$, the number of such paths originating from $w$ is at most $d^{k-1}$. Therefore, there are at most $t\ell(d-1)^{k-1}$ such vertices in $S_x\cap L_i$. \qedhere
  \end{enumerate}
\end{proof}

@@

\begin{proof}
  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.  Since $xy\in E(I)$, $G^{k,d}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.  Therefore $G$ contains a $(k,d)$ $vw$-shortcut $P$.  By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.  If neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.  By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.  By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.  By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.  But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
\end{proof}

@@

\begin{proof}
  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition of $H$. For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.  Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.  What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.

  Consider an arbitray node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r=x$.  To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.

@@

  Since $w\in S_{\delta}$, at least one of the following is true:
  \begin{enumerate}
    \item $w\in Y_\delta$; or 
    \item $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$.
  \end{enumerate}
  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{x_i}$.

  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.  For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.  Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-decreasing.

  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_p\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.  Therefore $s_p\ge\delta$.  To see that $s_p>\delta-1$,
  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.  By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p>\delta-1$.

  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$. Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.

  Next we observe that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H$.  Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.  By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.  Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.  By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.  Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.  Therefore $a_ia_i''\in E(H^+)$ as claimed.

  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.  By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$. 
  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t-1}{t}$.
\end{proof}

@@

\begin{thm}\thmlabel{modern-major-general}
  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t-1$.  Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(t\ell k^2d^k)$ where the treewidth of $H$ is at most $\binom{k+t-1}{t}-1$.
\end{thm}

\begin{proof}
  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{k,d}$.  In particular, $G^{k,d}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.  This increases the thickness given by \clmref{general-width} from $O(t\ell kd^k)$ to $O(t\ell k^2d^k)$.
\end{proof}

It is natural to ask if \thmref{modern-major-general} is tight.  We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.  If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ contains a clique $K$ of size  $s=2^{k/2+1}-1$.  If these vertices appear in $r$ parts of the $I$-partition, then the treewidth of $I$ is at least $r-1$.  On the other hand, these vertices must all appear on two consecutive layers, so the layered width of the partition is at least $s/(2r)$. Therefore any layered $I$ partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
\[
    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
\]   
\newpage

>>>>

For non-negative integers $k$ and $d$, a graph $G$ and distinct vertices $v,w\in V(G)$, a \emph{$(k,d)$ $vw$-shortcut} in $G$ is a path in $G$ with endpoints $v$ and $w$ of length at most $k$ and whose internal vertices each have degree at most $d$.  The \emph{$(k,d)^\text{th}$ power of $G$}, denoted $G^{k,d}$, is the graph with vertex $V(G^{k,d})=V(G)$ in which, for each distinct pair $v,w\in V(G)$, $vw\in E(G^{k,d})$ if and only if $G$ contains a $(k,d)$ $vw$-shortcut.

Let $T$ and, for each $x\in V(T)$, $Y_x$, $V_x$, $F_x$, and $N_x$ be a decomposition obtained by applying \thmref{generalized-tripod} to $G$, let $H:=G/\{Y_x:x\in V(T)\}$ denote the resulting quotient graph, and let $(B_x:x\in V(T))$ be a tree-decomposition of $H$ given by \lemref{nice-decomposition}.  For simplicity, we bypass the graph $H'$ and assume, from the outset, that $V(H)\subseteq V(T)$.

For two nodes $v,w\in V(G)$, a $(k,d)$ \emph{$vw$-longcut} is a walk $W=v_0,\ldots,v_{r}$ in $G$ of length $r\le k$ with $v_0=v$, $v_r=w$, in which, for each $i\in\{1,\ldots,r-1\}$, the degree of $v_i$ in $G$ is at most $d$.  Observe that every $(k,d)$ $vw$-shortcut is also a $(k,d)$ $vw$-longcut and that $G$ contains a $(k,d)$ $vw$-shortcut if and only if it contains a $(k,d)$ $vw$-longcut.

For a node $x\in V(T)$, we say that a $(k,d)$ $vw$-longcut $W$ in $G$ \emph{crosses} $x$ if $Y_x$ contains an internal vertex of $W$.  We say that a vertex $v\in V(G)$ \emph{participates} in $x$ if $v\in Y_x$ or $G$ contains a $vw$-longcut that crosses $x$.

\begin{clm}
  For every vertex $v\in V(G)$, there is a (unique) node $a(v)\in V(T)$ such that $v$ participates in $a(v)$ and $a(v)$ is a $T$-ancestor of every node $x$ in which $v$ participates. 
\end{clm}

@@

  Let $X_H:=\{x\in V(H): X\cap Y_x\neq\emptyset\}$.  The connectivity of $G[X]$ implies that $H[X_H]$ is connected.  
  %Property~\tref{ancestor-edge} implies that, for every $xy\in E(H)$, one of $x$ or $y$ is a $T$-ancestor of the other. 
  Choose $a(v)\in X_H$ to be any member of $X_H$ that does not have a strict $T$-ancestor in $X_H$.  Transitivity of the $T$-ancestor relationship, \tref{ancestor-edge}, and connectivity of $H[X_H]$ implies that $a(v)$ is a $T$-ancestor of every node $x\in X_H$, as required.
\end{proof}

For each $x\in V(T)$, we define the \emph{separator}
\[  
     S_x := \{v\in V(G): a(v)= x\} \enspace .
\]
Observe that $\mathcal{S}:=(S_x : x\in V(T))$ is a partition of $V(G)$.  We let $I:=G^{k,d}/\mathcal{S}$ denote the resulting quotient graph and we let $V(I)\subseteq V(T)$ in the obvious way, so that each $x\in V(I)$ is the vertex obtained by contracting $S_x$ in $G^{k,d}$. (Nodes $x\in V(T)$ with $S_x=\emptyset$ do not contribute a vertex to $I$.)

@@

  By \yref{y-subsets}, $Y_x\subseteq V_x$, so $v\not\in Y_x$.  Therefore, $G$ contains a path $v_0,v_1,\ldots,v_{r}$ such that $r <k$, $v_0=v$, $v_r\in Y_x$ and the degree of $v_i$ in $G$ is at most $d$ for each $i\in\{1,\ldots,r\}$.  Since $v\not\in V_x$ and $v_r\in V_x$, \yref{separator}, implies that $v_i\in F_x$ for some $i\in\{0,\ldots,r-1\}$. Now \yref{ancestor-edge} implies $v_i\in Y_a$ for some strict $T$-ancestor $a$ of $x$.  But the existence of the path $v_0,\ldots,v_i$ in $G$ implies that $v=v_0\in S_{a'}$ for some ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
\end{proof}

@@

\begin{clm}\clmlabel{general-width}
  For each $i\in\{1,\ldots,h\}$ and each $x\in V(I)$, $|S_x\cap V_i|\le (2k-1)\ell d^k$.  
\end{clm}

\begin{proof}
  Consider some vertex $v\in S_x\cap L_i$.  Since $v\in S_x$, $G$ contains a path $v_0,\ldots,v_r$ such that $r<k$, $v_0\in Y_x$, $v_r=v$, and the degree of $v_i$ in $G$ is at most $d$, for each $i\in\{0,\ldots,r-1\}$.  Note that $v_0$ is one of the at most $(2k-1)\ell$ vertices in $Z=\bigcup_{j=-k+1}^{k-1}(V_i\cap Y_x)$. For each $w\in Z$, the number of such paths originating from $w$ is at most $\sum_{r=0}^{k-1} d^{r}< d^k$ (for $d\ge 2$). Therefore, $|S_x\cap L_i|\le |Z|d^k\le (2k-1)\ell d^k$.
  % \end{enumerate}
\end{proof}

@@

\begin{proof}
  Suppose, for the sake of contradiction, that neither $x$ nor $y$ is a $T$-ancestor of the other.  Since $xy\in E(I)$, $G^{k,d}$ contains an edge $vw$ with $v\in S_x$ and $w\in S_y$.  Therefore $G$ contains a $(k,d)$ $vw$-shortcut $P$.  By \clmref{s-subset}, $v\in V_x$ and $w\in V_y$.  By \yref{containment-i}, if neither $x$ nor $y$ is a $T$-ancestor of the other, then $V_x$ and $V_y$ are disjoint.  By \yref{ancestor-edge} $N_x$ and $V_y$ are also disjoint.  By \yref{separator} $P$ contains an internal vertex $v'\in F_x$.  By \yref{ancestor-edge}, $v'\in Y_a$ for some strict $T$-ancestor $a$ of $x$.  But this implies that $v\in S_{a'}$ for some $T$-ancestor $a'$ of $a$, contradicting the assumption that $v\in S_x$.
\end{proof}

@@

\begin{proof}
  For the tree-decomposition $(C_x:x\in V(T))$ of $I$ we use the same tree $T$ used in the tree-decomposition $(B_x:x\in V(T))$ of $H$. For each node $x$ of $T$, $C_x$ contains $x$ as well as every $T$-ancestor $a$ of $x$ such that $I$ contains an edge $ax'$ where $x$ is a $T$-ancestor of $x'$ (including the possibility that $x=x'$).
  \clmref{i-ancestor} ensures that, for every edge $ax'\in E(I)$, $a,x'\in C_{x'}$ and the connectivity of $T[a]:=T[\{x\in V(T):a\in C_x]$ is obvious.  Therefore $(C_x:x\in V(T))$ is indeed a tree-decomposition of $I$.  What remains is to bound the size, $\max\{|C_x|:x\in V(T)\}$, of the largest bag.

  Consider an arbitrary node $x\in V(T)$ whose ancestors are $x_0,\ldots,x_{r-1},x_r$ where $x_0$ is the root of $T$ and $x_r:=x$.  To avoid triple-subscripts in what follows, we abuse notation slightly by using $V_i$, $F_i$, and $N_i$,  as shorthands for $V_{x_i}$, $F_{x_i}$ and $N_{x_i}$, respectively.

@@

  Since $w\in S_{\delta}$, at least one of the following is true:
  \begin{enumerate}
    \item $w\in Y_\delta$; or 
    \item $G$ contains a $(k,d)$ $ww'$-longcut $P_{ww'}$ that has an internal vertex in $Y_{\delta}$.
  \end{enumerate}
  In the former case we define $P_{ww'}$ to be the path of length 0 that contains only $w=w'$.  Let $w''$ denote the first vertex of $P_{ww'}$ contained in $Y_{\delta}$.

  Let $w_0,w_1,\ldots,w_p$ be the path that begins $w_0:=v'$ and then follows the subpath of $P_{ww'}$ that begins at $w_1:=w$ and ends at $w_p:=w''$.  For each $i\in\{0,\ldots,p\}$, let $s_i=\max\{j\in\{1,\ldots,r\}: \{w_0,\ldots,w_i\}\subseteq V_{j}\}\}$, and let $a_i=x_{s_i}$.  Note that $s_0,\ldots,s_p$ is a non-increasing sequence and $a_0,\ldots,a_p$ is a sequence of nodes of $T$ whose distance from the root is non-increasing.

  By definition, $a_0=x_r$.  We claim that $a_p=x_\delta$, i.e., $s_p=\delta$. 
  To see this, first observe that, for each $i\in\{1,\ldots,p\}$, $w_i\in V_{\delta}$ since, otherwise, an internal vertex of $P_{ww'}$ belong to $F_\delta$, which would imply (by \yref{ancestor-edge}) that $w\in S_{\delta'}$ for some $\delta' < \delta$, contradicting the assumption that $w\in S_\delta$.  Therefore $s_p\ge\delta$.  To see that $s_p<\delta-1$,
  observe that either $w=w''\in Y_\delta$ or $P_{ww'}$ contains an internal vertex $w''$ in $Y_\delta$.  By \yref{thickness} and the defnition of $V_x$, $V_{\delta-1}$ does not contain $w''$, so $s_p<\delta-1$.

  Let $H^+$ denote the supergraph of $H$ with vertex set $V(T)$ and in which $xy\in E(H^+)$ if and only there exists some $z\in V(T)$ such that $x,y\in B_z$. 
  % Note that, by \tref{ancestor-edge}, $xy\in E(H^+)$ if and only if $x\in B_y$ or $y\in B_x$.
  We claim that $a_0,\ldots,a_p$ is a lazy walk\footnote{A \emph{lazy walk} in a graph $H$ is a walk in the non-simple graph $H'$ obtained by adding a self loop to each vertex of $H$.} in $H^+$.  Indeed, if $a_i\neq a_{i+1}$ for some $i\in\{0,\ldots,p-1\}$ then this is precisely because $w_i\in V_{a_i}$ but $w_{i+1}\not\in V_{a_i}$.  By definition, $w_i\in Y_{a_i'}$ for some $T$-descendant $a_i'$ of $a_i$.
  By \yref{separator}, $w_{i+1}\in F_{a_i}$ so by \yref{ancestor-edge} $w_{i+1}\in Y_{a_i''}$ for some strict $T$-ancestor $a_i''$ of $a_i$.  Since $w_iw_{i+1}\in E(G)$, $a_i'a_i''\in E(H)$.  By \tref{subtree-root}, $a_i''\in B_{a_i''}$ and $a_i''\in B_{a_i'}$.  Since $a_i$ is on the path from $a_i'$ to $a_i''$ in $T$ this implies that $a_i''\in B_{a_i}$.  Therefore $a_ia_i''\in E(H^+)$ as claimed.

  Thus, $a_0,\ldots,a_p$ is a lazy walk in $H^+$ of length $p\le k$ where the distance $s_i$ between $a_i$ and the root $x_0$ of $T$ is non-decreasing.  By removing repeated vertices this gives a path in the directed graph $\overrightarrow{H}^+$ obtained by directing each edge $xy\in E(H^+)$ from its $T$-descendant $x$ towards its $T$-ancestor $y$. 
  Finally, we are in a position to appeal to \cite[Lemma~24]{pilipczuk.siebertz:polynomial-arxiv} which states that the number of nodes in $\overrightarrow{H}^+$ that can be reached from any node $x$ by a directed path of length at most $k$ is at most $\binom{k+t}{t}$.
\end{proof}

@@

\begin{thm}\thmlabel{modern-major-general}
  Let $\ell\ge 1$, $t\ge 2$, $d\ge 1$, and $k\ge 1$ be integers, let $G$ be a graph having a layered $H$-partition of layered width $\ell$ in which $H$ has treewidth at most $t-1$.  Then $G^{k,d}$ has a layered $I$-partition of layered width at most $O(k^2\ell d^k)$ where the treewidth of $H$ is at most $\binom{k+t}{t}-1$.
\end{thm}

\begin{proof}
  The proof is almost immediate, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{k,d}$.  In particular, $G^{k,d}$ may contain an edge $vw$ with $v\in V_i$ and $w\in V_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle V_0',\ldots,V_h'\rangle$ in which $V_i'=\bigcup_{j=ki}^{ki+k-1} V_i$.  This increases the thickness given by \clmref{general-width} from $O(k\ell d^k)$ to $O(k^2\ell d^k)$.
\end{proof}

It is natural to ask if \thmref{modern-major-general} is tight.  We do not answer this question, but we do point out that, even when $t=1$ and $d=3$, an exponential dependence on $k$ is necessary.  If $G$ is a complete binary tree on $2^{k/2}$ leaves, then $G^{k,3}$ is a clique of size  $s=2^{k/2+1}-1$.  Any partition of $G^{k,3}$ into $r$ parts will give an quotient graph $I$ that is an $r$-clique, which has treewidth $r-1$. On the other hand, any layering of $V(G^{k,3})$ has at most two layers, so the layered width of any partition into $r$ parts is at least $s/(2r)$. Therefore any layered $I$-partition of $G^{k,d}$ of layered width $\ell'$ in which the treewidth of $I$ is at most $t'-1$ must satisfy
\[
    \ell'\times t' \ge 2^{k/2}-1/2 \enspace .
\]   
\newpage
